### 2016-12-01 paper/paper.tex
init paper

\title{This is title}
\author{
        This is author information\\
        Name\\
        Email
}
\date{}
\documentclass[12pt]{article}
\usepackage[margin=0.7in]{geometry}
\usepackage{graphicx}
\usepackage{float}
\usepackage{comment}
\usepackage{amsmath}

\begin{document}
\maketitle

\begin{center}
\textbf{Abstract}
\end{center}
\noindent
This is Abstract section

\section{Introduction}
Introduction Section
\begin{enumerate}
\item Problem Statement
\end{enumerate}

\section{Background}
Background of Machine Learning and Deep Learning
\begin{enumerate}
\item Deep Learning\\

\item Deep Learning Library\\
	\begin{enumerate}
	\item DL4J\\
	\item Theano\\
	\item Tensorflow\\
	\end{enumerate}

\item Machine Learning in Driving Data\\
\end{enumerate}

\section{Data Set}
Explain data\\

\section{Technical Approach}
Theory and Method\\
implementation with Tensorflow\\

\section{Experiment Result}
Result and Visualization\\

\section{Conclusion and Future work}
Conclusion and Future work\\

\section*{Reference}
Reference Section\\

\end{document}